---
title: "Multiplex Canada Multiple Imputations"
author: "Fio Vialard"
Created: "August 21, 2023"
updates: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, fig.width=14, fig.height=14, message = F)

```

# Imputation plan

## Set-up

### Packages

```{r packages, message=FALSE}
require(mice)
require(lattice)
require(tidyverse)
require(gridExtra)
require(pan)
require(naniar)
require(flextable)
#install.packages("miceadds")
require(miceadds)
require(readxl)
require(lme4)
set.seed(123)
theme_set(theme_bw())
```

### Import dataset

```{r import data}
multiplex<- read.csv("3_intermediate/missing_data/20230908_multiplex_na_final.csv")
head(multiplex, 3)

source("2_scripts/00_functions.R")
```

### Factorize and remove baseline infection data

Since no NA and having the new infection and total infection data gives us enough information. Also remove the comments for now

```{r clean data}
multiplex <- multiplex %>% 
  select(!c(comments, baseline.hcv:baseline.syph)) %>%
 mutate(across(SexualActivity:Clinic.Location, as.factor)) %>% 
  rename(WorkStatus = WorkStatusStatus, Date.Recruited = DateRecuit)
summary(multiplex)
```

## Missing data pattern

```{r}
total.na <- sum(is.na(multiplex))

indiv.na<- sum(!complete.cases(multiplex))

prop.indiv.na <- round(indiv.na/nrow(multiplex) *100, digits =2)

multiplex.no.pro <- multiplex %>% 
  select(-c(FearPOC:PreferenceFollowup))

prop.indiv.na2 <- sum(!complete.cases(multiplex.no.pro))/nrow(multiplex.no.pro)*100
```

Total missing data: `r total.na`

Total number of participants with at least one missing data: `r indiv.na`

Percentage of participants with missing data: `r prop.indiv.na` %

Percentage of participants with missing data when patient-reported outcomes are removed: `r round(prop.indiv.na2, digits = 2)` %

```{r, fig.cap= "Pattern of missingness for variables in the Canada multiplex dataset. \nThe numbers on the left indicate the number of rows that have the same pattern of missingness. \nThe numbers on the right represent the number of missing data for each pattern. \nThe numbers at the bottom represent the total number of missing values for each variable with the total number of missing data in the dataset at the bottom right", purl=FALSE}

pattern<- md.pattern(multiplex, rotate.names = T)

na.patterns.total<- nrow(pattern)
```

Total numbers of patterns: `r na.patterns.total` . There is 1 individuals with `r as.integer(2/3*ncol(multiplex))` (or 2/3) missing variables.

### Which variables have similar patterns?

```{r, purl=FALSE}
na.fear <-sum(pattern[, "FearConventional"] == 0)

na.timing <- sum(pattern[, "PreferenceTiming"] == 0)

na.age <- sum(pattern[,"Age"]==0)

na.current.partners <- sum(pattern[, "CurrentPartner"]==0)

na.condom <- sum(pattern[,"CondomUse"]==0)

na.share.needles <- sum(pattern[,"ShareNeedles"]==0)
```

Patient-reported outcomes (e.g. fear of conventional strategy `r na.fear` and preference timing `r na.timing`) have similar missing patterns.

Socio-demographic outcomes from age (`r na.age`) to current partner (`r na.current.partners`) variable have the same pattern of missingness.

Risk factor outcomes from condom use (`r na.condom`) to sharing needles (`r na.share.needles`) variables have the same pattern of missingness.

### Remove observations with 2/3 or more missing variables

```{r remove observations with 2/3 missingness}
multiplex1 <- multiplex %>% 
  mutate(na_count = rowSums(is.na(multiplex))) %>%
  filter(na_count < 26) %>%
  select(!na_count)

# check 
nrow(multiplex1) == nrow(multiplex) - 1
```

## Association between variables

Are there any associations between variables. Let's find out by using two representative examples for each group of similar missing pattern.

Create columns for clinic, stbbi status and missingness column to be populated later

```{r, purl=FALSE}
multiplex4plots <- multiplex1 %>%
  mutate(site = ifelse(Clinic.Location == "REZO", "REZO", "RECAP"), 
         stbbi.pos = as.factor(ifelse(hcv.pos == 1| hiv.pos ==1 | syphilis.pos == 1, 
                                      "Infected","Non-infected")),
         missingness = NA)
```

### Does variable missingness depend on clinic?

```{r}
variable <- colnames(multiplex4plots)[3:21]

plots.site <-  get.correlation.site(multiplex4plots, variable, title = variable)
plots.site
```

### Does variable missingness depend on infection status?

7.  Fear of the conventional strategy

```{r, purl=FALSE}
plot7 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$FearConventional, 
                             title = "Fear of conventional strategy data")
```

8.  Timing preference

```{r, purl=FALSE}
plot8 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$PreferenceTiming, 
                             title = "Preference timing data")
```

9.  Age

```{r, purl=FALSE}
plot9 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$Age, 
                             title = "Age data")
```

10. Current partner

```{r, purl=FALSE}
plot10 <- get.correlation.inf(dataset = multiplex4plots, 
                              variable = multiplex4plots$CurrentPartner, 
                              title = "Current partner data")
```

11. sharing needles

```{r, purl=FALSE}
plot11 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$ShareNeedles, 
                             title = "Sharing needle data")
```

12. Condom use

```{r, purl=FALSE}
plot12 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$CondomUse, 
                             title = "Condom use data")
```

### Plots

```{r, purl=FALSE}
all.plots <- grid.arrange( plot7, 
                          plot8, plot9,  plot10,  plot11,  plot12, ncol=2)


ggsave(all.plots, filename = "3_intermediate/missing_data/missingness_correlation.png", 
       device = "png", width = 14, height = 14)
```

There are more missing data at RECAP and in non-infected individuals. This pattern is MNAR. Starting from Sharing needles, the missing data does not depend on site or infection status (because number of missing data is small).

# Multiple imputation methods

## Preparation

### Add site as a predictor

has to be integer for "2l.bin" method to work: 0 for REZO and 1 for RECAP

```{r transform data for imputation}
multiplex4imp <- multiplex1 %>% 
  mutate(site = as.integer(ifelse(Clinic.Location == "REZO", 0, 1))) %>% 
  select(-c(Clinic.Location))

summary(multiplex4imp$site)

summary(multiplex4imp)
```

### Initial imputation

Using default methods (5 iterations). No cluster

```{r, purl=FALSE}
ini.imp <- mice(multiplex4imp, nnet.MaxNWts = 2000, print = F)
# adjust the weights to not get an error message
# do not print the results
```

### Which variables were used as predictors?

Easier to visualize as a heatmap.

```{r, fig.cap= "Heatmap of predictor matrix. The columns are the variables predicted and rows show which value were acting as predictors. A value of 0 means no predicion occurred", purl=FALSE}
ini.pred <- ini.imp$predictorMatrix

prediction.matrix(ini.imp$predictorMatrix)
```

p_id and date were not predicted at all because all values were present in the dataset.

### What methods were used for each?

```{r, purl=FALSE}
ini.meth <- ini.imp$method
ini.meth
```

All used "polyreg" except for the variables without missing values and current partner used "logreg"

## 1-level imputation

### Proposed imputation plan

```{r, echo = F, purl=FALSE}
#create dataframe with each predictor
Parameters <- colnames(multiplex1)
categories <- list("Study ID, date and location", "Socio-demographic", "Clinical", "Test results", "Patient-reported")
Class <- sapply(multiplex1, class)

param <- data.frame(Parameters, Class, row.names = NULL)

param <- param %>% 
  mutate(class = ifelse(Parameters == "CurrentPartner", "binary", Class),
    Category = ifelse(
    Parameters == "p_id" |Parameters =="DateRecruited"|Parameters =="Clinic.Location", 
                                   categories[[1]], categories[[3]]),
  Category = ifelse(Class == "integer", categories[[4]], Category))

param[15:21,"Category"] <- categories[[2]]
param[22:33, "Category"] <- categories[[5]]

write.csv(param, "1_data/parameters.csv", row.names = F)
```

```{r, purl=FALSE}
param <- read_excel('1_data/parameters.xlsx')

param.1l <- param %>% 
  select(-c(`Proposed.imputation.method (2-level)`))
# table 
param.1l |>
  flextable(cwidth = 0.25) |> #turn into flextable object
  merge_v(j=1) |> #Merge cells in first column with same value
  theme_box() |> #Apply a theme for aesthetics
  autofit() |>#automatically set column widths to reasonable values
  fit_to_width(7.5) # size of the page
```

### Transformation

Transform the selected data to binary variables (see analysis plan)

```{r transform to binary variables}
multiplex4imp.1l <- multiplex4imp %>%
  mutate(
    SexualActivity = ifelse(
      SexualActivity %in% c("2-I do not wish to answer", "0-No"), 
      "0-No or no wish to answer", 
      as.character(SexualActivity)
    ),
    CondomUse = ifelse(
      CondomUse %in% c("1-Always", "2-Sometimes"), 
      "1-Ever", 
      as.character(CondomUse)
    ),
    SexualPartners = case_when(
      SexualPartners %in% c("0-0", "1-1", "2-2 to 5") ~ "0 to 5",
      SexualPartners %in% c("3-6 to 10", "4-11 or more") ~ "6 or more",
      TRUE ~ as.character(SexualPartners)
    ),
    PastSTD = ifelse(
     PastSTD  %in% c("2-I do not know", "0-No"), 
      "0-No or do not know", 
      as.character(PastSTD)
    ),
    PastInjectDrugs = ifelse(
      PastInjectDrugs %in% c("2-I do not wish to answer", "0-No"), 
      "0-No or no wish to answer", 
      as.character(PastInjectDrugs)
    ),
    ShareNeedles = ifelse(
      ShareNeedles %in% c("None", "0-No"),
      "0-No",
      as.character(ShareNeedles)
    ),
    Alcohol = ifelse(
      Alcohol %in% c("1-1-2 times per week", "2-3-5 times per week"),
      "1-5 times per week",
      as.character(Alcohol)
      ),
      # across(c(PastHIVTest, PastHCVTest, PastSyphilisTest), 
      #        ~ifelse(c(.) %in% c("2-Yes; 6 months to 1 year ago", 
    #"3-Yes; more than 1 year ago"),
      #                "more than 6 months ago", 
      #                as.character(.))) 
      # for now uncomment until we know how to deal with the variables
    PastSyphilis = ifelse(
     PastSyphilis %in% c("2-I do not know", "0-No"), 
      "0-No or do not know", 
      as.character(PastSyphilis)
    ),
    Age = case_when(
      Age %in% c("1-18-24", "1-25-34") ~ "18 to 34",
      Age %in% c("3-35-44", "4-45-54", "5-55 and above") ~ "35 and above",
      TRUE ~ as.character(Age)
    ),
    Gender = ifelse(
      Gender %in% "4-Bisexual",
      NA,
      as.character(Gender)
    ),
    Descent = case_when(
      Descent %in% c("4-European/ North American") 
      ~ "Non-Mediterranean European/North American",
      !Descent %in% c("Non-Mediterranean European/North American", NA) 
      ~ "Mediterranean European, African, Latin American, Caribbean and/or Asian",
      TRUE ~ as.character(Descent)
    ),
   EducationStatus= ifelse(
     !EducationStatus %in% c("4-Post-graduate degree", NA),
     "No post-graduate degree",
     as.character(EducationStatus)
   ),
   WorkStatus = case_when(
     WorkStatus %in% c("1-Employed (full time)", "2-Employed (part time)") 
     ~ "Currently employed",
     !WorkStatus %in% c("Currently employed", NA) ~ "Not currently employed"
   ),
  MonthlyIncomeStatus = ifelse(
    !MonthlyIncomeStatus %in% c("1-<$2;000 CAD", NA),
    "$2000 CAD and above",
    as.character(MonthlyIncomeStatus)
  ),
  PreferenceTiming = ifelse(
    !PreferenceTiming %in% c("0-1 day", NA),
    "2 days or more",
    as.character(PreferenceTiming)
  ),
  RecommendRapidTests = ifelse(
    RecommendRapidTests %in% c("0-No", "2-Not Sure"),
    "No or not sure", 
    as.character(RecommendRapidTests)
  ),
  PreferenceFollowup = case_when(
   PreferenceFollowup %in% c("3-Face-to-face", "4-No preference")
   ~ "No preference for digital follow-up",
   !PreferenceFollowup %in% c("No preference for digital follow-up", NA)
   ~"Digital follow-up",
   TRUE ~ as.character(PreferenceFollowup)
  )
      ) %>%
  mutate(across(SexualActivity:PreferenceFollowup, as.factor))

summary(multiplex4imp.1l)
```

### Run imputation with logistic regression for transformed factors

```{r, warning=FALSE, purl=FALSE}
# mice.1l <- mice(multiplex4imp.1l, nnet.MaxNWts = 2000, print = F, )
# 
# 
# summary(complete(mice.1l))
# 
# # save results 
# mice.1l.long<- complete(mice.1l, "long")
# 
# mice.1l.broad <- complete(mice.1l, "broad")
# 
# write.csv(mice.1l.long, "3_intermediate/imputed/imputed_data_long_1l.csv",
#          row.names = F)
# 
# write.csv(mice.1l.broad, "3_intermediate/imputed/imputed_data_broad_1l.csv",
#           row.names = F)
```

Not working for now, try to fix later

### Prediction matrix and methods

```{r, purl=FALSE}
# prediction.matrix(mice.1l$predictorMatrix)
# 
# meth.1l <- mice.1l$method
# meth.1l
```

The variables transformed into binary factors are now imputed using logistic regression

### Diagnostic plots

```{r, purl=FALSE}
# plot(mice.1l)
```

Variables with small amount of missing data (e.g. PastHIv/HCV/SyphilisTest) have unstable diagnostic plots.

## 1-level predictive mean matching

### Proposed imputation plan

```{r, purl=FALSE}
param.pmm <- param.1l %>% 
  mutate(Proposed.imputation.method = ifelse(
    Proposed.imputation.method %in% "None",
    "None",
    "Predictive mean matching"
  ))
# table 
param.pmm |>
  flextable(cwidth = 0.25) |> #turn into flextable object
  merge_v(j=1) |> #Merge cells in first column with same value
  theme_box() |> #Apply a theme for aesthetics
  autofit() |>#automatically set column widths to reasonable values
  fit_to_width(7.5)
```

```{r, purl=FALSE}
mice.pmm <- mice(multiplex4imp.1l, method = "pmm", nnet.MaxNWts = 2000, print = F)

summary(complete(mice.pmm))

# save 
mice.pmm.long<- complete(mice.pmm, "long")

mice.pmm.broad <- complete(mice.pmm, "broad")

write.csv(mice.pmm.long, "3_intermediate/imputed/imputed_data_long_pmm.csv",
         row.names = F)

write.csv(mice.pmm.broad, "3_intermediate/imputed/imputed_data_broad_pmm.csv",
          row.names = F)

```

### Prediction matrix and methods

```{r, purl=FALSE}
prediction.matrix(mice.pmm$predictorMatrix)
meth.pmm <- mice.pmm$method 
meth.pmm
```

### Diagnostic plots

```{r, purl=FALSE}
plot(mice.pmm)
```

## 2-level imputation

### Proposed imputation method

```{r, purl=FALSE}
param.2l <- param %>% 
  select(-c(Proposed.imputation.method))
# table 
param.2l |>
  flextable(cwidth = 0.25) |> #turn into flextable object
  merge_v(j=1) |> #Merge cells in first column with same value
  theme_box() |> #Apply a theme for aesthetics
  autofit() |>#automatically set column widths to reasonable values
  fit_to_width(7.5)
```

As described in: [Van Buuren, S. (2018) Flexible Imputation of Missing Data (2nd ed).](https://stefvanbuuren.name/fimd/sec-mlguidelines.html)

### Transformation

For this method to work, the variables have to be transformed to numbers

```{r transform to numbers, , purl=FALSE}
multiplex4imp.2l <- multiplex4imp.1l %>%
  mutate(across(c(3:33), ~as.factor(case_when(
    . %in% c("0-No or no wish to answer", "0-Never", "0-No or do not know", "None", "0-No", 
             "No post-graduate degree", "Not currently employed", "1-<$2;000 CAD", 
             "0-Very satisfied", "2-No preference","No preference for digital follow-up",
             "No or not sure") ~0,
    . %in% c("1-Yes", "1-Ever", "0 to 5", "0-HIV", "Yes", "1-5 times per week", 
             "1-Yes; less than 6 months ago", "18 to 34", "1-Male", 
             "Mediterranean European, African, Latin American, Caribbean and/or Asian", 
             "4-Post-graduate degree", "Currently employed", "$2000 CAD and above", 
             "0-1-no fear","0-1-no pain", "0-1 - no discomfort", "1-Satisfied", "0-1 day", 
             "1-Phlebotomy", "Digital follow-up") ~1,
    . %in% c("6 or more", "2-Herpes simplex", "2-Yes; 6 months to 1 year ago", 
             "35 and above", "2-Female", "Non-Mediterranean European/North American", "1-2",
             "2-Neutral/ not sure", "2 days or more", "0-Rapid test") ~2,
    . %in% c("3-Chlamydia", "3-Yes; more than 1 year ago", "3-Transgender", "2-3",
             "3-Dissatisfied")~3,
    . %in% c("4-Gonorrhoeae", "5-No response", "3-4" )~4,
    . %in% c("5-Syphilis", "4-5-highest")~5,
    . %in% c("6-Other")~6,
    T~NA))))
   
summary(multiplex4imp.2l)
```

### Change prediction matrix

to include 2-level pmm for selected variables (see analysis plan)

```{r predictor matrix, , purl=FALSE}
# Create the predictor matrix
pred.2l <- make.predictorMatrix(multiplex4imp.2l)

# set site as cluster variables 
pred.2l[c(4, 15:33), "site"] <- -2

#check
prediction.matrix(pred.2l)
```

### Adapt methods

For now we will keep variables not predicted by cluster variables as the initial method

```{r methods, , purl=FALSE}
# run quick imputation
ini.quick <- mice(multiplex4imp.2l, maxit = 0)

# copy initial methods
meth.2l<- ini.quick$method

#replace selected variables to 2l_pmm
meth.2l[c(4, 15:33)] <- "2l.pmm"

# check 
meth.2l
```

### Re-run imputation

```{r imputation, warning= FALSE, message = F, , purl=FALSE}

# mice.2l <- mice(multiplex4imp.2l, nnet.MaxNWts = 2000, pred= pred.2l, meth = meth.2l, print =F)

```

also not working

### View imputed data

```{r, purl=FALSE}
# summary(complete(mice.2l))
# 
# summary(multiplex4imp.2l)
```

### Extract imputed data and save as csv

```{r, message=FALSE, purl=FALSE}
# mice.2l.long<- complete(mice.2l, "long")
# 
# mice.2l.broad <- complete(mice.2l, "broad")
# 
# write.csv(mice.2l.long, "3_intermediate/imputed/imputed_data_long_2l.csv",
#          row.names = F)
# 
# write.csv(mice.2l.broad, "3_intermediate/imputed/imputed_data_broad_2l.csv",
#           row.names = F)

```

### Diagnostic plot

With 5 iteration

```{r, purl=FALSE}
# plot(mice.2l)
```

with 20 (using the same object and running 15 more iterations)

```{r, purl=FALSE}
# mice.2l.20 <- mice.mids(mice.2l, nnet.MaxNWts = 2000, print =F, maxit=15)
# 
# plot(mice.2l.20)
```

## Sensitivity analysis

### Density plot of selected variables 1l imputation

```{r fig.cap="density plot of the values selected observed variables (blue) and their corresponding imputed values (red).", message=FALSE, purl=FALSE}

# dplot1 <- densityplot(mice.1l, ~SexualActivity)
# dplot2 <- densityplot(mice.1l, ~CondomUse)
# dplot3 <- densityplot(mice.1l, ~SexualPartners)
# dplot4 <- densityplot(mice.1l, ~PastSTD)
# dplot5 <- densityplot(mice.1l, ~TypePastSTD) 
# dplot6 <- densityplot(mice.1l, ~PastInjectDrugs)
# dplot7 <- densityplot(mice.1l, ~ShareNeedles)
# dplot8 <- densityplot(mice.1l, ~Alcohol) 
# dplot9 <- densityplot(mice.1l, ~PastHIVTest) 
# dplot10 <- densityplot(mice.1l, ~PastSyphilisTest) 
# 
# dplots.clin <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
#                        dplot7, dplot8, dplot9, dplot10, ncol =2 )
# dplots.clin
# 
# ggsave (dplots.clin, filename = "3_intermediate/imputed/Densityplots_observed_vs_1l-imputed_data_clinical.png", 
#          device = "png", width = 14, height = 14)
# 
# dplot1 <- densityplot(mice.1l, ~Age)
# dplot2 <- densityplot(mice.1l, ~Gender)
# dplot3 <- densityplot(mice.1l, ~Descent)
# dplot4 <- densityplot(mice.1l, ~EducationStatus)
# dplot5 <- densityplot(mice.1l, ~WorkStatus)
# dplot6 <- densityplot(mice.1l, ~MonthlyIncomeStatus)
# dplot7 <- densityplot(mice.1l, ~CurrentPartner)
# 
# dplots.ses <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
#                        dplot7, ncol =2 )
# 
# ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_1l-imputed_data_ses.png", 
#          device = "png", width = 14, height = 14)
# 
# dplot1 <- densityplot(mice.1l, ~DiscomfortPOC) 
# dplot2 <- densityplot(mice.1l, ~FearConventional)
# dplot3 <- densityplot(mice.1l, ~RetestPreference)
# dplot4 <- densityplot(mice.1l, ~SatisfactionConventional)
# dplot5 <- densityplot(mice.1l, ~PreferenceTiming)
# dplot6 <- densityplot(mice.1l, ~PainPOC)
# 
# dplot.pro <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6)
# 
# dplot.pro
# 
# ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_1l-imputed_data_pro.png", 
#          device = "png", width = 14, height = 14)
```

For some variables, the imputed data follows the shape of the observed data.

### Density plot of selected variables 1l imputation - pmm

```{r, message=FALSE, purl=FALSE}
dplot1 <- densityplot(mice.pmm, ~SexualActivity)
dplot2 <- densityplot(mice.pmm, ~CondomUse)
dplot3 <- densityplot(mice.pmm, ~SexualPartners)
dplot4 <- densityplot(mice.pmm, ~PastSTD)
dplot5 <- densityplot(mice.pmm, ~TypePastSTD) 
dplot6 <- densityplot(mice.pmm, ~PastInjectDrugs)
dplot7 <- densityplot(mice.pmm, ~ShareNeedles)
dplot8 <- densityplot(mice.pmm, ~Alcohol) 
dplot9 <- densityplot(mice.pmm, ~PastHIVTest) 
dplot10 <- densityplot(mice.pmm, ~PastSyphilisTest) 

dplots.clin <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
                       dplot7, dplot8, dplot9, dplot10, ncol =2 )
dplots.clin

ggsave (dplots.clin, filename = "3_intermediate/imputed/Densityplots_observed_vs_pmm-imputed_data_clinical.png", 
         device = "png", width = 14, height = 14)

dplot1 <- densityplot(mice.pmm, ~Age)
dplot2 <- densityplot(mice.pmm, ~Gender)
dplot3 <- densityplot(mice.pmm, ~Descent)
dplot4 <- densityplot(mice.pmm, ~EducationStatus)
dplot5 <- densityplot(mice.pmm, ~WorkStatus)
dplot6 <- densityplot(mice.pmm, ~MonthlyIncomeStatus)
dplot7 <- densityplot(mice.pmm, ~CurrentPartner)

dplots.ses <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
                       dplot7, ncol =2 )

ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_pmm-imputed_data_ses.png", 
         device = "png", width = 14, height = 14)

dplot1 <- densityplot(mice.pmm, ~DiscomfortPOC) 
dplot2 <- densityplot(mice.pmm, ~FearConventional)
dplot3 <- densityplot(mice.pmm, ~RetestPreference)
dplot4 <- densityplot(mice.pmm, ~SatisfactionConventional)
dplot5 <- densityplot(mice.pmm, ~PreferenceTiming)
dplot6 <- densityplot(mice.pmm, ~PainPOC)

dplot.pro <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6)

dplot.pro

ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_pmm-imputed_data_pro.png", 
         device = "png", width = 14, height = 14)
```

pmm predicts the missing data to be closer to the predicted data

### Density plot of selected variables pmm2-level imputation

```{r message=FALSE, purl=FALSE}
# dplot1 <- densityplot(mice.2l, ~SexualActivity)
# dplot2 <- densityplot(mice.2l, ~CondomUse)
# dplot3 <- densityplot(mice.2l, ~SexualPartners)
# dplot4 <- densityplot(mice.2l, ~PastSTD)
# dplot5 <- densityplot(mice.2l, ~TypePastSTD) 
# dplot6 <- densityplot(mice.2l, ~PastInjectDrugs)
# dplot7 <- densityplot(mice.2l, ~ShareNeedles)
# dplot8 <- densityplot(mice.2l, ~Alcohol) 
# dplot9 <- densityplot(mice.2l, ~PastHIVTest) 
# dplot10 <- densityplot(mice.2l, ~PastSyphilisTest) 
# 
# dplots.clin <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
#                        dplot7, dplot8, dplot9, dplot10, ncol =2 )
# dplots.clin
# 
# ggsave (dplots.clin, filename = "3_intermediate/imputed/Densityplots_observed_vs_2l-imputed_data_clinical.png", 
#          device = "png", width = 14, height = 14)
# 
# dplot1 <- densityplot(mice.2l, ~Age)
# dplot2 <- densityplot(mice.2l, ~Gender)
# dplot3 <- densityplot(mice.2l, ~Descent)
# dplot4 <- densityplot(mice.2l, ~EducationStatus)
# dplot5 <- densityplot(mice.2l, ~WorkStatus)
# dplot6 <- densityplot(mice.2l, ~MonthlyIncomeStatus)
# dplot7 <- densityplot(mice.2l, ~CurrentPartner)
# 
# dplots.ses <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
#                        dplot7, ncol =2 )
# 
# ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_2l-imputed_data_ses.png", 
#          device = "png", width = 14, height = 14)
# 
# dplot1 <- densityplot(mice.2l, ~DiscomfortPOC) 
# dplot2 <- densityplot(mice.2l, ~FearConventional)
# dplot3 <- densityplot(mice.2l, ~RetestPreference)
# dplot4 <- densityplot(mice.2l, ~SatisfactionConventional)
# dplot5 <- densityplot(mice.2l, ~PreferenceTiming)
# dplot6 <- densityplot(mice.2l, ~PainPOC)
# 
# dplot.pro <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6)
# 
# dplot.pro
# 
# ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_2l-imputed_data_pro.png", 
#          device = "png", width = 14, height = 14)
```

Prediction is even better.

## Random intercept mixed model regression

### Original data

We will conduct a random intercept model multiple log regression with HCV as the dependent variable, site as the cluster variable and selected variables as independent variables :

$y_i = \beta_0 + \beta_1X_{ij} + u_j$ with $u_j\sim N(0, \sigma_j^2)$ - random intercept at site level

1.  SexualActivity
2.  PastSTD
3.  PastInjectDrugs
4.  Age
5.  Gender

```{r, purl=FALSE}

fit.original <- glm(hcv.pos~
                      SexualActivity+PastSTD+PastInjectDrugs+Age+Gender+ (1:site),
                    data = multiplex4imp.1l)

summary(fit.original) 
```

Some values have huge standard errors. This probably corresponds to variables with a lot of missing data or levels with little amounts of data (e.g. transgender). But the binary trasnformation got rid of most of them.

### Imputed dataset

```{r, purl=FALSE}
# fit.imp <- with(mice.1l,
#             glm(hcv.pos~
#                       SexualActivity+PastSTD+PastInjectDrugs+Age+Gender+(1:site)))
# 
# # pool the data
# pool.fit.imp <- pool(fit.imp)
# 
# summary(pool.fit.imp, conf.int = T)
```

### Pmm results

```{r, purl=FALSE}
fit.imp <- with(mice.pmm,
            glm(hcv.pos~
                      SexualActivity+PastSTD+PastInjectDrugs+Age+Gender+ (1:site)))

# pool the data
pool.fit.imp <- pool(fit.imp)

summary(pool.fit.imp, conf.int = T)
```

### 2l-results

```{r, purl=FALSE}
# fit.imp <- with(mice.2l,
#             glm(hcv.pos~
#                       SexualActivity+PastSTD+PastInjectDrugs+Age+Gender+ (1:site)))
# 
# # pool the data
# pool.fit.imp <- pool(fit.imp)
# 
# summary(pool.fit.imp, conf.int = T)
```

Results are all very similar. Which method is best?

# Updates

Added code for dataset that excludes PROMs

Keep 1-level no PROMs for projpred for now

```{r, import dataset 2}
multiplex_no_prom <- read.csv("3_intermediate/missing_data/multiplex_na_final_no_prom.csv")
head(multiplex_no_prom, 3)

multiplex_no_prom <- multiplex_no_prom %>% 
  select(!c(baseline.hcv:baseline.syph)) %>%
 mutate(across(SexualActivity:Clinic.Location, as.factor)) %>% 
  rename(WorkStatus = WorkStatusStatus, Date.Recruited = DateRecuit)
summary(multiplex_no_prom)
```

```{r, fig.cap= "Pattern of missingness for variables in the Canada multiplex dataset (without PROM). \nThe numbers on the left indicate the number of rows that have the same pattern of missingness. \nThe numbers on the right represent the number of missing data for each pattern. \nThe numbers at the bottom represent the total number of missing values for each variable with the total number of missing data in the dataset at the bottom right", purl=FALSE}

pattern<- md.pattern(multiplex_no_prom, rotate.names = T)

na.patterns.total<- nrow(pattern)
```

There are a total of `r na.patterns.total` patterns

```{r, remove row with more than 2/3 NA}

large.na<- as.integer(2/3*ncol(multiplex_no_prom))

multiplex_no_prom <- multiplex_no_prom %>% 
  mutate(na_count = rowSums(is.na(multiplex_no_prom))) %>%
  filter(na_count < large.na) %>%
  select(!na_count)

```

Create columns for clinic, stbbi status and missingness column to be populated later

```{r, purl=FALSE}
multiplex4plots <- multiplex_no_prom %>%
  mutate(site = ifelse(Clinic.Location == "REZO", "REZO", "RECAP"), 
         stbbi.pos = as.factor(ifelse(hcv.pos == 1| hiv.pos ==1 | syphilis.pos == 1, 
                                      "Infected","Non-infected")),
         missingness = NA)
```

### Does variable missingness depend on clinic?

same as above but without the PROMs.

Use get correlation function to get all plots at once

```{r}
variable <- colnames(multiplex4plots)[3:21]

plots <-  get.correlation.site(multiplex4plots, variable, title = variable)

```

### Does variable missingness depend on infection status?

9.  Age

```{r, purl=FALSE}
plot9 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$Age, 
                             title = "Age data")
```

10. Current partner

```{r, purl=FALSE}
plot10 <- get.correlation.inf(dataset = multiplex4plots, 
                              variable = multiplex4plots$CurrentPartner, 
                              title = "Current partner data")
```

11. sharing needles

```{r, purl=FALSE}
plot11 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$ShareNeedles, 
                             title = "Sharing needle data")
```

12. Condom use

```{r, purl=FALSE}
plot12 <- get.correlation.inf(dataset = multiplex4plots, 
                             variable = multiplex4plots$CondomUse, 
                             title = "Condom use data")
```

### Plots

```{r, purl=FALSE}
plots.per.site <- grid.arrange(plots[[1]], plots[[2]], plots[[3]],
                               plots[[4]], plots[[5]], plots[[6]],
                               plots[[7]], plots[[8]], plots[[9]],
                               plots[[10]], plots[[11]], plots[[12]],
                               plots[[13]], plots[[14]], plots[[15]],
                               plots[[16]], plots[[17]], plots[[18]],
                               plots[[19]],
                               ncol=3)

plots.per.status <- grid.arrange(plot9, plot10, plot11, plot12, ncol= 2)

ggsave(plots.per.status, filename = "3_intermediate/missing_data/missingness_no_prom.png", device = png)

ggsave(plots.per.site, filename = "4_outputs/missingness_site.png", 
       device = "png", width = 14, height = 14)
```

There are more missing data at RECAP and in non-infected individuals. This pattern is MNAR. Starting from Sharing needles, the missing data does not depend on site or infection status (because number of missing data is small).

## Imputation

```{r  transform data for imputation 2}
multiplex4imp2 <- multiplex_no_prom %>% 
  mutate(site = as.integer(ifelse(Clinic.Location == "REZO", 0, 1))) %>% 
  select(-c(Clinic.Location))

summary(multiplex4imp2$site)

summary(multiplex4imp2)
```

sing default methods (5 iterations). No cluster

```{r, purl=FALSE}
ini.imp <- mice(multiplex4imp2, nnet.MaxNWts = 2000, print = F)
# adjust the weights to not get an error message
# do not print the results
```

### Which variables were used as predictors?

```{r, fig.cap= "Heatmap of predictor matrix. The columns are the variables predicted and rows show which value were acting as predictors. A value of 0 means no predicion occurred", purl=FALSE}
ini.pred <- ini.imp$predictorMatrix

prediction.matrix(ini.imp$predictorMatrix)
```

### What methods were used for each?

```{r, purl=FALSE}
ini.meth <- ini.imp$method
ini.meth
```

## Transformation

Transform the selected data to binary variables (see analysis plan)

```{r transform to binary variables 2}
multiplex4imp.1l2 <- multiplex4imp2 %>%
  mutate(
    SexualActivity = ifelse(
      SexualActivity %in% c("2-I do not wish to answer", "0-No"), 
      "0-No or no wish to answer", 
      as.character(SexualActivity)
    ),
    CondomUse = ifelse(
      CondomUse %in% c("1-Always", "2-Sometimes"), 
      "1-Ever", 
      as.character(CondomUse)
    ),
    SexualPartners = case_when(
      SexualPartners %in% c("0-0", "1-1", "2-2 to 5") ~ "0 to 5",
      SexualPartners %in% c("3-6 to 10", "4-11 or more") ~ "6 or more",
      TRUE ~ as.character(SexualPartners)
    ),
    PastSTD = ifelse(
     PastSTD  %in% c("2-I do not know", "0-No"), 
      "0-No or do not know", 
      as.character(PastSTD)
    ),
    PastInjectDrugs = case_when(
      PastInjectDrugs %in% c("2-I do not wish to answer", "0-No")~ 
      "0-No or no wish to answer",
      PastInjectDrugs %in% "Yes"~ "1-Yes",
      T~as.character(PastInjectDrugs)
    ),
    ShareNeedles = ifelse(
      ShareNeedles %in% c("None", "0-No"),
      "0-No",
      as.character(ShareNeedles)
    ),
    Alcohol = ifelse(
      Alcohol %in% c("1-1-2 times per week", "2-3-5 times per week"),
      "1-5 times per week",
      as.character(Alcohol)
      ),
      # across(c(PastHIVTest, PastHCVTest, PastSyphilisTest), 
      #        ~ifelse(c(.) %in% c("2-Yes; 6 months to 1 year ago", 
    #"3-Yes; more than 1 year ago"),
      #                "more than 6 months ago", 
      #                as.character(.))) 
      # for now uncomment until we know how to deal with the variables
    PastSyphilis = ifelse(
     PastSyphilis %in% c("2-I do not know", "0-No"), 
      "0-No or do not know", 
      as.character(PastSyphilis)
    ),
    Age = case_when(
      Age %in% c("1-18-24", "1-25-34") ~ "18 to 34",
      Age %in% c("3-35-44", "4-45-54", "5-55 and above") ~ "35 and above",
      TRUE ~ as.character(Age)
    ),
    Gender = ifelse(
      Gender %in% "4-Bisexual",
      NA,
      as.character(Gender)
    ),
    Descent = case_when(
      Descent %in% c("4-European/ North American") 
      ~ "Non-Mediterranean European/North American",
      !Descent %in% c("Non-Mediterranean European/North American", NA) 
      ~ "Mediterranean European, African, Latin American, Caribbean and/or Asian",
      TRUE ~ as.character(Descent)
    ),
   EducationStatus= ifelse(
     !EducationStatus %in% c("4-Post-graduate degree", NA),
     "No post-graduate degree",
     as.character(EducationStatus)
   ),
   WorkStatus = case_when(
     WorkStatus %in% c("1-Employed (full time)", "2-Employed (part time)") 
     ~ "Currently employed",
     !WorkStatus %in% c("Currently employed", NA) ~ "Not currently employed"
   ),
  MonthlyIncomeStatus = ifelse(
    !MonthlyIncomeStatus %in% c("1-<$2;000 CAD", NA),
    "$2000 CAD and above",
    as.character(MonthlyIncomeStatus)
  # ),
  # PreferenceTiming = ifelse(
  #   !PreferenceTiming %in% c("0-1 day", NA),
  #   "2 days or more",
  #   as.character(PreferenceTiming)
  # ),
  # RecommendRapidTests = ifelse(
  #   RecommendRapidTests %in% c("0-No", "2-Not Sure"),
  #   "No or not sure", 
  #   as.character(RecommendRapidTests)
  # ),
  # PreferenceFollowup = case_when(
  #  PreferenceFollowup %in% c("3-Face-to-face", "4-No preference")
  #  ~ "No preference for digital follow-up",
  #  !PreferenceFollowup %in% c("No preference for digital follow-up", NA)
  #  ~"Digital follow-up",
  #  TRUE ~ as.character(PreferenceFollowup)
  )
      ) %>%
  mutate(across(SexualActivity:CurrentPartner, as.factor))

summary(multiplex4imp.1l2)
```

### Run imputation with logistic regression for transformed factors

```{r, warning=FALSE}
mice.1l <- mice(multiplex4imp.1l2, nnet.MaxNWts = 2000, print = F)

summary(complete(mice.1l))

# save results 
mice.1l.long<- complete(mice.1l, "long", include=TRUE)

mice.1l.broad <- complete(mice.1l, "broad", include = T)

write.csv(mice.1l.long, "3_intermediate/imputed/imputed_data_long_no_prom_1l.csv", row.names = F)

write.csv(mice.1l.broad, "3_intermediate/imputed/imputed_data_broad_no_prom_1l.csv")
```

### Prediction matrix and methods

```{r, purl=FALSE}
prediction.matrix(mice.1l$predictorMatrix)

meth.1l <- mice.1l$method
meth.1l
```

The variables transformed into binary factors are now imputed using logistic regression

### Diagnostic plots

```{r, purl=FALSE}
plot(mice.1l)
```

### add more

```{r, purl = FALSE}
mice.1l.20 <- mice.mids(mice.1l, nnet.MaxNWts = 2000, print =F, maxit=15)

plot(mice.1l.20)
```

## Sensitivity analysis

### Density plot of selected variables 1l imputation

```{r fig.cap="density plot of the values selected observed variables (blue) and their corresponding imputed values (red).", message=FALSE, purl=FALSE}

dplot1 <- densityplot(mice.1l, ~SexualActivity)
  
dplot2 <- densityplot(mice.1l, ~CondomUse)
dplot3 <- densityplot(mice.1l, ~SexualPartners)
dplot4 <- densityplot(mice.1l, ~PastSTD)
dplot5 <- densityplot(mice.1l, ~TypePastSTD) 
dplot6 <- densityplot(mice.1l, ~PastInjectDrugs)
dplot7 <- densityplot(mice.1l, ~ShareNeedles)
dplot8 <- densityplot(mice.1l, ~Alcohol) 
#dplot9 <- densityplot(mice.1l, ~PastHIVTest) 
dplot10 <- densityplot(mice.1l, ~PastSyphilisTest)

dplots.clin <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
                       dplot7, dplot8, dplot10, ncol =2 )
dplots.clin

ggsave (dplots.clin, filename = "3_intermediate/imputed/Densityplots_observed_vs_1l-imputed_data_clinical_no_prom.png", 
         device = "png", width = 14, height = 14)

dplot11 <- densityplot(mice.1l, ~Age)
dplot12 <- densityplot(mice.1l, ~Gender)
dplot13 <- densityplot(mice.1l, ~Descent)
dplot14 <- densityplot(mice.1l, ~EducationStatus)
dplot15 <- densityplot(mice.1l, ~WorkStatus)
dplot16 <- densityplot(mice.1l, ~MonthlyIncomeStatus)
dplot17 <- densityplot(mice.1l, ~CurrentPartner)

dplots.ses <- grid.arrange(dplot11, dplot12, dplot13, dplot14, dplot15, dplot16,
                       dplot17, ncol =2)

ggsave (dplots.ses, filename = "3_intermediate/imputed/Densityplots_observed_vs_1l-imputed_data_ses_no_prom.png", 
         device = "png", width = 14, height = 14)

all.dplots <- grid.arrange(dplot1, dplot2, dplot3, dplot4, dplot5, dplot6,
                       dplot7, dplot8, dplot10, dplot11, dplot12, dplot13, dplot14, dplot15, dplot16,
                       dplot17, ncol = 2)

ggsave(all.dplots, filename = "4_outputs/density_1l_imp.png",
       device = "png", width = 14, height = 20)
```

# References

<https://www.rdocumentation.org/packages/mice/versions/3.16.0>

**See vignettes for examples**
